---
title: "Quant Management Assignment#6"
output:
  html_document: default
  pdf_document: default
---

#This notebook contains the code for the Assignment 6.

#Install lpSolveAPI package if not alresdy installed
```{r}
#install.packages("lpSolveAPI")
```

# Load the library
```{r}
library(lpSolveAPI)
```

# Solution 1 using inequalities
# Let us set up the Linear problem. Note that we have 6 decision variables and 5 constraints. 
```{r}
lprec <- make.lp(5, 6)
```

# Set the minimization objective function
```{r, results='hide'}
set.objfn(lprec, c(622, 614, 630, 641, 645, 649))
lp.control(lprec,sense='min')
```

# Set values for the rows (set the Left hand side constraints)
```{r}
set.row(lprec, 1, c(1, 1, 1), indices = c(1, 2, 3))
set.row(lprec, 2, c(1, 1, 1), indices = c(4, 5, 6))
set.row(lprec, 3, c(1, 1), indices = c(1, 4))
set.row(lprec, 4, c(1, 1), indices = c(2, 5))
set.row(lprec, 5, c(1, 1), indices = c(3, 6))
```

# Set the right hand side values
```{r}
rhs <- c(100, 120, 80, 60, 70)
set.rhs(lprec, rhs)
```

# Set constraint type and set variable types and bound
```{r}
set.constr.type(lprec, c("<=", "<=", "=", "=", "="))
set.bounds(lprec, lower = rep(0, 6))
```

# Finally, name the decision variables (column) and constraints (rows)
```{r}
lp.rownames <- c("CapacityA", "CapacityB", "DemandW1", "DemandW2", "DemandW3")
lp.colnames <- c("PlantAW1", "PlantAW2", "PlantAW3","PlantBW1", "PlantBW2", "PlantBW3")
dimnames(lprec) <- list(lp.rownames, lp.colnames)
```

# View the linear problem object to make sure it's correct
```{r}
lprec
```

# Save this into a file
```{r}
write.lp(lprec, filename = "AED.lp", type = "lp")
```

# Now, solve the model
```{r}
solve(lprec)
```

# Show the value of objective function, variables, constraints and slack
```{r}
get.objective(lprec)

get.variables(lprec)

get.constraints(lprec)

get.constraints(lprec) - rhs
```


***
# Solution 2: Using Dummy Variables 
# Let us set up the Linear problem. Note that we had 8 decision variables and 6 constraints. 
```{r}
lprec1 <- make.lp(6, 8)
```

# Set the minimization objective function
```{r, results='hide'}
set.objfn(lprec1, c(622, 614, 630, 0, 641, 645, 649, 0))
lp.control(lprec1, sense='min')
```

# Set values for the rows (set the Left hand side constraints)
```{r}
set.row(lprec1, 1, c(1, 1, 1, 1), indices = c(1, 2, 3, 4))
set.row(lprec1, 2, c(1, 1, 1, 1), indices = c(5, 6, 7, 8))
set.row(lprec1, 3, c(1, 1), indices = c(1, 5))
set.row(lprec1, 4, c(1, 1), indices = c(2, 6))
set.row(lprec1, 5, c(1, 1), indices = c(3, 7))
set.row(lprec1, 6, c(1, 1), indices = c(4, 8))
```

# Set the right hand side values
```{r}
rhs <- c(100, 120, 80, 60, 70, 10)
set.rhs(lprec1, rhs)
```

# Set constraint type and set variable types and bound
```{r}
set.constr.type(lprec1, c("=", "=", "=", "=", "=", "="))
set.bounds(lprec1, lower = rep(0, 8))
```

# Finally, name the decision variables (column) and constraints (rows)
```{r}
lp.rownames <- c("CapacityA", "CapacityB", "DemandW1", "DemandW2", "DemandW3", "DemandW4")
lp.colnames <- c("PlantAW1", "PlantAW2", "PlantAW3", "PlanntAW4", "PlantBW1", "PlantBW2", "PlantBW3", "PlantBW4")
dimnames(lprec1) <- list(lp.rownames, lp.colnames)
```

# View the linear program object to make sure it's correct
```{r}
lprec1
```

# Now solve the model
```{r}
solve(lprec1)
```

# Show the value of objective function, variables, constraints and slack
```{r}
get.objective(lprec1)

get.variables(lprec1)

get.constraints(lprec1)

get.constraints(lprec1) - rhs
```

***
# Also,We can now read the lp formulation using an lp file and solve it. I am using the same lp file which I have saved above.

# Read from file and solve it
```{r}
x <- read.lp("AED.lp")     # create an lp object x
x                          # display x

solve(x)                   # Solution
get.objective(x)           # get objective value
get.variables(x)           # get values of decision variables
get.constraints(x)         # get constraints
```
