---
title: "Quantitative Management Assignment#8"
author: "Rakhee Moolchandani"
date: "11/01/2020"
output: 
  pdf_document: default
  html_document: default
---

# This notebook contains the code for the Assignment 8.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using Benchmarking Libraries for DEA

## We will now run DEA analysis using the benchmarking library. First, install the library, if you don't have it already.
```{r,message=FALSE}
#install.packages("Benchmarking")
#install.packages("readxl")
library(Benchmarking)
library(readxl)
```

## Now, we read our input data. We will read the data from an excel file. Remember our problem had 6 DMUs with two inputs and two outputs.
## Inputs: Staffing Labor, Cost of Supplies
## Outputs: No of patient-days reimbursed by third party, No of patient-days reimbursed privately 
```{r}
#Read the data from excel file
data <- read_excel("DEA.xlsx")
#See the data
data
#Facility1 to Facility 6 are the DMUs
namesDMU <- data[1]
namesDMU
#Lets see the Inputs 
inputs <- data[c(2,3)]
inputs
#Now, see the outputs
outputs <- data[c(4,5)]
outputs
#Create the input matrix
x <- matrix(c(data$`Staff Hours per Day`,data$`Supplies per Day`),ncol = 2)
#Lets see the input matrix
x
#Create the output matrix
y <- matrix(c(data$`Reimbursed Patient-Days`,data$`Privately Paid Patient-Days`),ncol = 2)
#Lets see the output matrix
y
#plot the graph for Inputs
dea.plot.isoquant(x[,1],x[,2])
#plot the graph for Outputs
dea.plot.transform(y[,1],y[,2])
```

## We now run the DEA analysis for different assumptions:

## We use the option of FDH, Free disposability hull, no convexity assumption 
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
e1 <- dea(x,y,RTS = "FDH")
#Show the Efficiency
e1
#Show the list of objects calculated
str(e1)
#Show the peers
peers(e1)
#Show the lambda
lambda(e1)
#Add the Efficiency, Peers & Lambda values in the table
report1 <- cbind(data, e1$eff, e1$lambda, e1$peers)
#Name the columns of the table
colnames(report1)<- c(names(namesDMU),names(inputs), names(outputs),'Efficiency','Lambda1','Lambda2','Lambda3','Lambda4','Lambda5','Lambda6','Peers')
#Sow the table
report1
#plot the graph for FDH Assumption
dea.plot(x,y,RTS="FDH",txt = rownames(report1))
```


## The results indicate that DMUs 1, 2, 3 , 4, 5 and 6 all are efficient.


## We use the option of CRS, Constant Return to Scale, convexity and free disposability 
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
e2 <- dea(x,y,RTS = "CRS")
#Show the Efficiency
e2
#Show the list of objects calculated
str(e2)
#Show the peers
peers(e2)
#Show the lambda
lambda(e2)
#Add the Efficiency & Lambda values in the table
report2 <- cbind(data, e2$eff, e2$lambda)
#Name the columns of the table
colnames(report2)<- c(names(namesDMU),names(inputs), names(outputs),'Efficiency','Lambda1','Lambda2','Lambda3','Lambda4','Lambda5','Lambda6')
#Sow the table
report2
#plot the graph for CRS Assumption
dea.plot(x,y,RTS="CRS",txt = rownames(report2))
```


## The results indicate that DMUs 1, 2, 3 and 4 are efficient. DMU(5) is only 97.7% efficient, and DMU(6) is 86.7% efficient. 


## We use the option of VRS, Variable returns to scale, convexity and free disposability
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
e3 <- dea(x,y,RTS = "VRS")
#Show the Efficiency
e3
#Show the list of objects calculated
str(e3)
#Show the peers
peers(e3)
#Show the lambda
lambda(e3)
#Add the Efficiency & Lambda values in the table
report3 <- cbind(data, e3$eff, e3$lambda)
#Name the columns of the table
colnames(report3)<- c(names(namesDMU),names(inputs), names(outputs),'Efficiency','Lambda1','Lambda2','Lambda3','Lambda4','Lambda5','Lambda6')
#Show the table
report3
#plot the graph for VRS Assumption
dea.plot(x,y,RTS="VRS",txt = rownames(report3))
```


## The results indicate that DMUs 1, 2, 3, 4 and 5 are efficient. DMU(6) is only 89.6% efficient. 


## We use the option of IRS, Increasing returns to scale, convexity and free disposability 
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
e4 <- dea(x,y,RTS = "IRS")
#Show the Efficiency
e4
#Show the list of objects calculated
str(e4)
#Show the peers
peers(e4)
#Show the lambda
lambda(e4)
#Add the Efficiency & Lambda values in the table
report4 <- cbind(data, e4$eff, e4$lambda)
#Name the columns of the table
colnames(report4)<- c(names(namesDMU),names(inputs), names(outputs),'Efficiency','Lambda1','Lambda2','Lambda3','Lambda4','Lambda5','Lambda6')
#Show the table
report4
#plot the graph for IRS Assumption
dea.plot(x,y,RTS="IRS",txt = rownames(report4))
```


## The results indicate that DMUs 1, 2, 3, 4 and 5 are efficient. DMU(6) is only 89.6% efficient. 


## We use the option of DRS, Decreasing returns to scale, convexity, down-scaling and free disposability
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
e5 <- dea(x,y,RTS = "DRS")
#Show the Efficiency
e5
#Show the list of objects calculated
str(e5)
#Show the peers
peers(e5)
#Show the lambda
lambda(e5)
#Add the Efficiency, Peers & Lambda values in the table
report5 <- cbind(data, e5$eff, e5$lambda)
#Name the columns of the table
colnames(report5)<- c(names(namesDMU),names(inputs), names(outputs),'Efficiency','Lambda1','Lambda2','Lambda3','Lambda4','Lambda5','Lambda6')
#Show the table
report5
#plot the graph for IRS Assumption
dea.plot(x,y,RTS="DRS",txt = rownames(report5))
```


## The results indicate that DMUs 1, 2, 3 and 4 are efficient. DMU(5) is only 97.7% efficient, and DMU(6) is 86.7% efficient. 



## We use the option of FRH, Additivity (scaling up and down, but only with integers), and free disposability 
```{r}
#DEA input or output efficiency measures, peers, lambdas and slacks
e6 <- dea(x,y,RTS = "ADD")
#Show the Efficiency
e6
#Show the list of objects calculated
str(e6)
#Show the peers
peers(e6)
#Show the lambda
lambda(e6)
#Add the Efficiency, Peers & Lambda values in the table
report6 <- cbind(data, e6$eff, e6$lambda)
#Name the columns of the table
colnames(report6)<- c(names(namesDMU),names(inputs), names(outputs),'Efficiency','Lambda1','Lambda2','Lambda3','Lambda4','Lambda5','Lambda6')
#Sow the table
report6
#plot the graph for FDH Assumption
dea.plot(x,y,RTS="ADD",txt = rownames(report5))
```


## The results indicate that DMUs 1, 2, 3 , 4, 5 and 6 all are efficient.


## Compare and Contrast the above Results
## Lets compare the efficiency of all the DMUs for all the assumptions
```{r}
#Add the Efficiency of all the DMUs for all the Assumptions in a table
EfficiencyReport <- cbind(data[,1],e1$eff,e2$eff,e3$eff,e4$eff,e5$eff,e6$eff)
#Name the columns of the table
colnames(EfficiencyReport) <- c(names(namesDMU),'FDH Efficiency','CRS Efficiency','VRS Efficiency','IRS Efficiency','DRS Efficiency','FRH Efficiency')
#Show the Efficiency table
EfficiencyReport
#Let's compare the Efficiency of all the DMUs for all the assumptions using a plot
#Concatenate the Efficiency
spreadsheet <- cbind(e1$eff,e2$eff,e3$eff,e4$eff,e5$eff,e6$eff) 
#Name the rows
rownames(spreadsheet) <- c("Facility1","Facility2","Facility3","Facility4","Facility5","Facility6")
#Name the columns
colnames(spreadsheet) <- c ("FDH","CRS","VRS","IRS","DRS","FRH")
#See the result
spreadsheet
#plot the graph
barplot(t(spreadsheet),col=cm.colors(6),xlab = "DMUs", ylab="Efficiency",beside=TRUE)
```

## 1) Facility 1 is fully efficient for all the assumptions.
## 2) Facility 2 is fully efficient for all the assumptions.
## 3) Facility 3 is fully efficient for all the assumptions.
## 4) Facility 4 is fully efficient for all the assumptions.
## 5) Facility 5 is fully efficient for FDH, VRS, IRS and FRH assumptions. For assumptions DRS and CRS, it is 97.7% efficient.
## 6) Facility 6 is fully efficient for FDH and FRS assumptions. For CRS and DRS assumptions, it is 86.7% efficient. For IRS and VRS assumptions, it is 89.6% efficient.




