---
title: "Quant Management Assignment#11"
author: "Rakhee Moolchandani"
date: "11/22/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

***

# Integer Programming Problem:

AP is a shipping service that guarantees overnight delivery of packages in the continental US. The company has various hubs at major cities and airports across the country. Packages are received at hubs, and then shipped to intermediate hubs or to their final destination.
The manager of the AP hub in Cleveland is concerned about labor costs, and is interested in determining the most effective way to schedule workers. The hub operates seven days a week, and the number of packages it handles varies from one day to another. The table below provides an estimate of the number of workers needed each day of the week.

Day of the Week | Workers Required
----------------|------------------
Sunday          |     18
Monday          |     27
Tuesday         |     22
Wednesday       |     26
Thursday        |     25
Friday          |     21
Saturday        |     19
                
Package handlers at AP are guaranteed a five-day work week with two consecutive days off. The base wage for the handlers is $750 per week. Workers working on Saturday or Sunday receive an additional $25 per day. The possible shifts and salaries for package handlers are:

Shift | Days off              | Wage
------|-----------------------|------------------
1     | Sunday and Monday     | $755  
2     | Monday and Tuesday    | $800
3     | Tuesday and Wednesday | $800     
4     | Wednesday and Thursday| $800    
5     | Thursday and Friday   | $800    
6     | Friday and Saturday   | $775    
7     | Saturday and Sunday   | $750

## Questions:
The manager wants to keep the total wage expenses as low as possible while ensuring that there are sufficient number of workers available each day. Formulate and solve the problem. What was the total cost? How many workers are available each day?

## Answers:
### Lets Formulate the problem 

Let us consider the decision variables as following:   
x1= the number of workers assigned to shift 1  
x2= the number of workers assigned to shift 2  
x3= the number of workers assigned to shift 3  
x4= the number of workers assigned to shift 4  
x5= the number of workers assigned to shift 5  
x6= the number of workers assigned to shift 6  
x7= the number of workers assigned to shift 7  

Minimize the total wage expense (Objective Function):  
Min = 775x1 + 800x2 + 800x3 + 800x4 + 800x5 + 775x6 + 750x7  

Workers Required each day (Constraints): 

Day of Week| Workers                                        
-----------|----------------------------------------------
Sunday     | 0x1 + 1x2 + 1x3 + 1x4 + 1x5 + 1x6 + 0x7 >= 18
Monday     | 0x1 + 0x2 + 1x3 + 1x4 + 1x5 + 1x6 + 1x7 >= 27
Tuesday    | 1x1 + 0x2 + 0x3 + 1x4 + 1x5 + 1x6 + 0x7 >= 22
Wednesday  | 1x1 + 1x2 + 0x3 + 0x4 + 1x5 + 1x6 + 1x7 >= 26
Thursday   | 1x1 + 1x2 + 1x3 + 0x4 + 0x5 + 1x6 + 1x7 >= 25
Friday     | 1x1 + 1x2 + 1x3 + 1x4 + 0x5 + 0x6 + 1x7 >= 21
Saturday   | 1x1 + 1x2 + 1x3 + 1x4 + 1x5 + 0x6 + 0x7 >= 19

x1, x2, x3, x4, x5, x6, x7 >=0

### Lets solve the IP problem, 
```{r}
# Load the library
library(lpSolveAPI) 
# Note that we had 7 decision variables and 7 constraints. 
lprec <- make.lp(7, 7)
# Set the minimization objective function
set.objfn(lprec, c(775, 800, 800, 800, 800, 775, 750))
lp.control(lprec,sense='min')
# Set values for the rows (set the Left hand side constraints)
set.row(lprec, 1, c(0, 1, 1, 1, 1, 1, 0), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(lprec, 2, c(0, 0, 1, 1, 1, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(lprec, 3, c(1, 0, 0, 1, 1, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(lprec, 4, c(1, 1, 0, 0, 1, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(lprec, 5, c(1, 1, 1, 0, 0, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(lprec, 6, c(1, 1, 1, 1, 0, 0, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(lprec, 7, c(1, 1, 1, 1, 1, 0, 0), indices = c(1, 2, 3, 4, 5, 6, 7))
# Set the right hand side values
rhs <- c(18, 27, 22, 26, 25, 21, 19)
set.rhs(lprec, rhs)
# Set constraint type and set variable types and bound
set.constr.type(lprec, c(">=", ">=", ">=", ">=", ">=", ">=", ">="))
# Set lower bound as zero
set.bounds(lprec, lower = rep(0, 7))
# Set variable type as integer
set.type(lprec,1:7,"integer")
# Finally, name the decision variables (column) and constraints (rows)
lp.rownames <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
lp.colnames <- c("Shift1", "Shift2", "Shift3","Shift4", "Shift5", "Shift6","Shift7")
dimnames(lprec) <- list(lp.rownames, lp.colnames)
# View the linear program object to make sure it's correct
lprec
# Save this into a file
write.lp(lprec, filename = "QA11.lp", type = "lp")
# Now solve the model
solve(lprec)
# Show the value of objective function, variables and constraints 
get.objective(lprec)
get.variables(lprec)
get.constraints(lprec)
```

### Also, We now read the lp formulation using an lp file saved above
```{r}
# Read the IP formulation file
y <- read.lp("QA11.lp")
# Print the model
y
# Solve the model
solve(y)
# Get the objective function value
get.objective(y)
```

### The total cost is $25675  


```{r}
# Get the variables value
get.variables(y)
D <- data.table::data.table(Decision_Variables=lp.rownames<-c("x1","x2","x3","x4","x5","x6","x7"),values=get.variables(y))
D
# Get the constraints value
get.constraints(y)
C <- data.table::data.table(Days_of_week=lp.rownames<-c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"),Workers_available=get.constraints(y))
C
```

### The above table shows the number of workers available each day

***