---
title: "Quant Management Assignment#4"
output:
  html_document: default
  pdf_document: default
---

#This notebook contains the code for the Assignment 4.

#Install lpSolveAPI package if not alresdy installed
```{r}
#install.packages("lpSolveAPI")
```

#Now, load the library
```{r}
library(lpSolveAPI)
```

#Let us set up the Weigelt Corporation problem. Note that we had 9 decision variables, and 11 constraints. 
```{r}
lprec <- make.lp(11, 9)
```

# set the maximization objective function
```{r}
set.objfn(lprec, c(420, 360, 300, 420, 360, 300, 420, 360, 300))
lp.control(lprec,sense='max')
```

# set values for the rows (set the Left hand side constraints)
```{r}
set.row(lprec, 1, c(1, 1, 1), indices = c(1, 2, 3))
set.row(lprec, 2, c(1, 1, 1), indices = c(4, 5, 6))
set.row(lprec, 3, c(1, 1, 1), indices = c(7, 8, 9))
set.row(lprec, 4, c(20, 15, 12), indices = c(1, 2, 3))
set.row(lprec, 5, c(20, 15, 12), indices = c(4, 5, 6))
set.row(lprec, 6, c(20, 15, 12), indices = c(7, 8, 9))
set.row(lprec, 7, c(1, 1, 1), indices = c(1, 4, 7))
set.row(lprec, 8, c(1, 1, 1), indices = c(2, 5, 8))
set.row(lprec, 9, c(1, 1, 1), indices = c(3, 6, 9))
set.row(lprec, 10, c(900, 900, 900, -750, -750 ,-750), indices = c(1, 2, 3, 4, 5, 6))
set.row(lprec, 11, c(450, 450, 450, -900, -900, -900), indices = c(4, 5, 6, 7, 8, 9))
```

# set the right hand side values
```{r}
rhs <- c(700, 900, 450, 13000, 12000, 5000, 900, 1200, 750, 0, 0)
set.rhs(lprec, rhs)
```

# set constraint type and set variable types and bound
```{r}
set.constr.type(lprec, c("<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "=", "="))
set.bounds(lprec, lower = rep(0, 9))
```

# finally, name the decision variables (column) and constraints (rows)
```{r}
lp.rownames <- c("Plant1Production", "Plant2Production", "Plant3Production", "Plant1Space", "Plant2Space", "Plant3Space", "Plant1Forecast", "Plant2Forecast", "Plant3Forecast", "Capacity1", "Capacity2")
lp.colnames <- c("Plant1Large", "Plant1Medium", "Plant1Small","Plant2Large", "Plant2Medium", "Plant2Small","Plant3Large", "Plant3Medium", "Plant3Small")
dimnames(lprec) <- list(lp.rownames, lp.colnames)
```

# view the linear program object to make sure it's correct
```{r}
lprec
```

#Save this into a file
```{r}
write.lp(lprec, filename = "weiglet.lp", type = "lp")
```

# now solve the model
```{r}
solve(lprec)
```

#Show the value of objective function, variables, constraints and slack
```{r}
get.objective(lprec)

get.variables(lprec)

get.constraints(lprec)

get.constraints(lprec) - rhs
```

***
#Also,We can now read the lp formulation using an lp file and solve it. I am using the same lp file which I have saved above.

#read from file and solve it
```{r}
x <- read.lp("weiglet.lp") # create an lp object x
x                          # display x

solve(x)                   # Solution
get.objective(x)           # get objective value
get.variables(x)           # get values of decision variables
get.constraints(x)         # get constraints
```
